---
title: "Introduction for the UKgrid Dataset"
author: "Rami Krispin (@Rami_Krispin)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### The UK National Electricity Transmission System Dataset

#### Intro

The `UKgrid` is an R dataset package with the historical national demand of the electricity transmission system in the UK and other related variables. This dataset is a half-hourly time series data with observations since January 2011. This dataset was sourced from National Grid UK [website](https://www.nationalgrid.com/uk).

#### Installation

Install the package from [CRAN](https://CRAN.R-project.org/package=UKgrid):

```{r eval=FALSE, message=TRUE, warning=TRUE, include=TRUE}
install.packages("UKgrid")
```

or install the development version from [Github](https://github.com/RamiKrispin/UKgrid):

```{r eval=FALSE, message=TRUE, warning=TRUE, include=TRUE}
# install.packages("devtools")
devtools::install_github("RamiKrispin/UKgrid")
```

### Usage

```{r}
library(UKgrid)

data("UKgrid")

str(UKgrid)
```

A variable dictionary is available in the dataset documentation. 

#### The `extract_grid` function

The `extract_grid` function provides the ability to extract the UKgrid series in a different format (`xts`, `zoo`, `ts`, `data.frame`, `data.table` and `tbl`), frequencies (half-hourly, hourly, daily, weekly, monthly and quarterly), and subset the series by time frame.   

For example, you can select the national demand variable (ND), using `xts` format:

```{r fig.height=5, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}

nd_half_hourly <- extract_grid(type = "xts", # default
                          columns = "ND", # default
                          aggregate = NULL # default
                          )

library(TSstudio)

ts_info(nd_half_hourly)

ts_plot(ts.obj = nd_half_hourly, 
        title = "UK National Demand - Half-Hourly")

```

Alternatively, you can aggregate the series to an hourly frequency with the `aggregate` argument:

```{r fig.height=5, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}

nd_hourly <- extract_grid(type = "ts", 
                          columns = "ND", 
                          aggregate = "hourly" 
                          )

class(nd_hourly)
frequency(nd_hourly)

ts_plot(ts.obj = nd_hourly, 
        title = "UK National Demand - Hourly")

```


Selection of the UKgrid columns is done by the `columns` argument. The full list of columns is available on the dataset documentation (`?UKgrid`). For instance, let's select the "ND" and "TSD" columns in a daily format:

```{r fig.height=5, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}

df <- extract_grid(type = "tbl", 
                          columns = c("ND","TSD"), 
                          aggregate = "daily" 
                          )

head(df)

ts_plot(ts.obj = df, 
        title = "UK National and Transmission System Demand - Daily")

```

Note: by default, when any of the data frame family structure is used, the output will include the timestamp of the data (even if was not selected in the columns argument)

Last but not least, you can subset the series by time range with the `start` and `end` argument:

```{r fig.height=5, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}

df1 <- extract_grid(type = "tbl", 
                          columns = "ND", 
                          aggregate = "daily", 
                          start = 2015,
                          end = 2017)

head(df1)

ts_plot(ts.obj = df1, 
        title = "UK National and Transmission System Demand - Daily between 2015 and 2017")

```
